Temporal Variations In Vapor Intrusion-Induced Indoor Air Contaminant Concentrations

### 2019-04-15 tex/main.tex
Last changes before submission

% TODO: Add abstract
\begin{abstract}

\end{abstract}

@@

The ASU house data were obtained over a period of several years.
During part of this time, controlled pressure method (CPM) tests were being conducted, in which the house was underpressurized to an extent greater than that characterizing “normal” house operation thus increasing VI potential\cite{mchugh_evaluation_2012,mchugh_recent_2017,holton_evaluation_2015}.
The period of CPM testing is thus excluded from the analysis.
Likewise, the existence of a preferential pathway at the ASU house needs to be considered in examining that dataset; during some of the testing at that site, this pathway was cut off, resulting in “normal” VI conditions in which the main source of contaminant was diffusion of contaminant vapor from an underlying groundwater source.\par

@@

The ground surface and the pipe are both sources of air to the soil domain.
Both are assumed to exist be at reference atmospheric pressure.
Soil gas transport is governed by Richard’s equation, a modified version of Darcy’s Law, taking the variability of soil moisture in the vadose zone into account\cite{richards_capillary_1931}.
The van Genuchten equations are used to predict the soil moisture content and thus the effective permeability of the soil\cite{van_genuchten_closed-form_1980}.
The effective diffusivity of contaminant in soil is calculated using the Millington-Quirk model\cite{millington_permeability_1961}.
The transport of contaminant vapor in the soil is assumed to be governed by the advection-diffusion equation, in which either advection or diffusion may dominate depending upon position and particular circumstances.
The key working equations and the boundary conditions are summarized in Table \ref{tbl:eqns_bc_parameters}.\par

@@

\begin{figure}[htb!]
 \centering
 \caption{Mean values of the maximum change in indoor air contaminant concentration that may be expected
  over a given time period. (e.g., 1D is 1 day, 2W is 2 weeks, and 3M is 3 months). The error bars are the 95\% confidence intervals.}\label{fig:resampling}
 \includegraphics[width=\textwidth]{temporal_variability/resampling.pdf}
\end{figure}

High frequency measurement of indoor air contaminant concentrations, $c_\mathrm{in}$, such as those in Figure \ref{fig:indianapolis}, took place at both the ASU House and the Indianapolis House over significant periods (Indianapolis: ca 1.7 years, ASU house: ca 3.5 years)\cite{u.s._environmental_protection_agency_assessment_2015,holton_temporal_2013}.
Furthermore, at the Indianapolis site the $c_\mathrm{in}$ for three different contaminants, chloroform, TCE, and tetrachloroethylene (PCE) were all collected, allowing examination of the variability of each VI contaminant.
The NAS North Island NAS dataset was obtained over a much shorter duration (9 days), and is therefore not examined in this portion of the analysis.
It should also be noted that the ASU house used 4-hour sorbent tubes, while Indianapolis took instantaneous "grab" samples.\par

@@

@@

We examine the relationship between $p_\mathrm{in/out}$ and $c_\mathrm{in}$ by constructing the two-dimensional kernel density estimation (KDE) plots seen in Figure \ref{fig:kde}.
The KDE plots allow us to view the measured distributions of $p_\mathrm{in/out}$ and $c_\mathrm{in}$, and develop a visual impression of how well these distributions correlate with one another.
For this analysis we considered two VI sites, NAS North Island and the ASU House.
The ASU House dataset was divided into two periods, one before and the other after the land drain (called the preferential pathway (PP) from  ere on) had been closed.
By comparing these two periods on a single plot, the impact of the preferential pathway becomes clearer.\par

@@

The above was proven by a second simulation, where the model was rerun with the preferential pathway present, but with the permeable (gravel) layer in the subslab removed and replaced by the surrounding soil (sandy loam).
This gave a ”uniform soil” scenario the results of which are shown as an orange line in the top panel of Figure \ref{fig:land_drain_scenarios}.
This simulation demonstrates that without a permeable subslab to effectively allow the ”advective potential” to be realized, existence of preferential pathway way actually not impact a VI site very much.
In order for a preferential pathway to significantly contribute to VI, this requires a scenario involving good advective communication between it the indoor environment.
These requirements were met at the ASU House.\par

@@

The model is also able to capture the weak trend in $\alpha_\mathrm{gw}$ with $p_\mathrm{in/out}$ when a preferential pathway is absent, when there exists a permeable subslab region.
These results are shown in the bottom panel of Figure \ref{fig:land_drain_scenarios}.  
These results are again in agreement with what was observed at the ASU House when the preferential pathway was closed, i.e. that there was a much more modest variation in indoor air concentration, irrespective of pressure, when the preferential pathway was cut off.\par

@@

% TODO: Simplify or clarify the figure, in particular how you use the legend to highlight the max change.
\begin{figure}
 \caption{Transient simulation of a "typical" VI day, using diurnal indoor/outdoor pressure difference and air exchange rate as inputs. Effect of preferential pathway considered.}\label{fig:diurnal}
 \includegraphics[width=\textwidth]{diurnal.pdf}
\end{figure}

@@

The result of these simulations are shown in the bottom two panels of Figure \ref{fig:diurnal}, where the left and right panels show the results of open and closed preferential pathways, respectively.
The "max change" value in the legends is the quotient of the lowest and highest predicted concentrations, i.e. a value of two indicate that the maximum daily concentration is twice as high as the lowest.
This quantity may be compared with the value that is plotted for "one day" in Figure \ref{fig:resampling}.
When the preferential pathway is open, there is a maximum daily variation of roughly a factor of 5, irrespective of if $A_e$ fluctuates or not, which is somewhat more than the maximum daily variation shown in Figure \ref{fig:resampling}.
The relatively small difference between the variable and constant $A_e$ cases indicates that most of the variability during a "typical" day is here attributable to fluctuations in $p_\mathrm{in/out}$, i.e. the contaminant transport into the modeled structure is advection dominated.
Even for the small fluctuations in $p_\mathrm{in/out}$ the contaminant entry rate fluctuation drives the observed indoor concentration.
When the preferential pathway is closed the story is quite different.
When air exchange rate is held constant, there is essentially no variation in $c_\mathrm{in}$.
This is again not surprising, as Figure \ref{fig:land_drain_scenarios} demonstrated that when the preferential pathway is closed, the influence of $p_\mathrm{in/out}$ on contaminant entry rate (and subsequently $c_\mathrm{in}$) is small.
Combined with the small $p_\mathrm{in/out}$ show that the contaminant transport into the modeled structure in this scenario is dominated by diffusion.
When the air exchange rate is allowed to fluctuate the maximum daily variation in $c_\mathrm{in}$ is 1.68, which is in line with what is shown in Figure \ref{fig:resampling}.
This shows that for a "typical" day, when the preferential pathway is closed off, much of the daily variation in $c_\mathrm{in}$ is due to daily fluctuations in air exchange rate.\par

These results demonstrate the complicated nature of temporal variability in $c_\mathrm{in}$.
It is important to recall that only the effects of indoor/outdoor pressure difference and air exchange rate have been considered here, but slower processes, e.g. changes in groundwater contaminant concentration or various seasonal effects can also have a significant impact on VI over time.
For the shorter time periods of concern in recent studies of temporal variability in indoor contaminant concentrations we believe that these are dominated by combinations of indoor/outdoor pressure differentials and air exchange rate.
For a site where advective communication between the subsurface and the indoor is good, $p_\mathrm{in/out}$ is likely a significant determinant of $c_\mathrm{in}$ and its temporal variability.
We have shown that that such a scenario may arise due to a preferential pathway, entering a permeable sub-base, but may also exist even in the absence of a preferential pathway just as the results from NAS North Island demonstrate.
At sites where advective transport into the structure is limited, much of the temporal variability in $c_\mathrm{in}$ may be attributed to natural fluctuations in air exchange rate.\par

>>>>

\begin{abstract}
Temporal variability in indoor air contaminant concentrations at vapor intrusion (VI) sites has been a concern for some time.
We consider the source of the reported variability at VI sites located near Hill Air Force Base (AFB) in Utah, an EPA experimental house in Indiana, and Naval Air Station North Island in California.
We focus in particular on how the indoor/outdoor pressure differences and air exchange rates affected indoor air contaminant concentrations at these sites.
We investigate how these dynamics differ for a site that is characterized by a preferential pathway (like Hill AFB) and VI sites that are not influenced by such pathways, using three-dimensional fluid dynamics models and statistical analysis of the aforementioned sites.
For a preferential pathway to impact a VI site, there must exist a medium allowing effective communication between a contaminant-delivering preferential pathway and the indoor air space, e.g. a permeable subslab space that may be provided by a gravel layer.
At sites characterized by significant advective transport from the subslab to the indoor air space, much of the short-term variability in indoor air contaminant concentration can be explained by an impact of fluctuations in indoor/outdoor pressure differences.
Meanwhile, air exchange rate variation drives most of the short-term variability at sites characterized by minor variations in advective transport.
\end{abstract}

@@

The ASU house data were obtained over a period of several years.
During part of this time, controlled pressure method (CPM) tests were being conducted, in which the house was underpressurized to an extent greater than that characterizing “normal” house operation, thus increasing VI potential\cite{mchugh_evaluation_2012,mchugh_recent_2017,holton_evaluation_2015}.
The period of CPM testing is thus excluded from the analysis.
Likewise, the existence of a preferential pathway at the ASU house needs to be considered in examining that dataset; during some of the testing at that site, this pathway was cut off, resulting in “normal” VI conditions in which the main source of contaminant was diffusion of contaminant vapor from an underlying groundwater source.\par

@@

The ground surface and the pipe are both sources of air to the soil domain.
Both are assumed to exist at reference atmospheric pressure.
Soil gas transport is governed by Richard’s equation, a modified version of Darcy’s Law, taking the variability of soil moisture in the vadose zone into account\cite{richards_capillary_1931}.
The van Genuchten equations are used to predict the soil moisture content and thus the effective permeability of the soil\cite{van_genuchten_closed-form_1980}.
The effective diffusivity of contaminant in soil is calculated using the Millington-Quirk model\cite{millington_permeability_1961}.
The transport of contaminant vapor in the soil is assumed to be governed by the advection-diffusion equation, in which either advection or diffusion may dominate depending upon position and particular circumstances.
The key working equations and the boundary conditions are summarized in Table \ref{tbl:eqns_bc_parameters}.\par

@@

High frequency measurement of indoor air contaminant concentrations, $c_\mathrm{in}$, such as those in Figure \ref{fig:indianapolis}, took place at both the ASU House and the Indianapolis House over significant periods (Indianapolis: ca 1.7 years, ASU house: ca 3.5 years)\cite{u.s._environmental_protection_agency_assessment_2015,holton_temporal_2013}.
Furthermore, at the Indianapolis site $c_\mathrm{in}$ for three different contaminants, chloroform, TCE, and tetrachloroethylene (PCE) were all collected, allowing examination of the variability of each VI contaminant.
The NAS North Island NAS dataset was obtained over a much shorter duration (9 days), and is therefore not examined in this portion of the analysis.
It should also be noted that the ASU house used 4-hour sorbent tubes, while Indianapolis took instantaneous "grab" samples.\par

@@

\begin{figure}[htb!]
 \centering
 \caption{Mean values of the maximum change in indoor air contaminant concentration that may be expected
  over a given time period. (e.g., 1D is 1 day, 2W is 2 weeks, and 3M is 3 months). The error bars are the 95\% confidence intervals.}\label{fig:resampling}
 \includegraphics[width=\textwidth]{resampling.pdf}
\end{figure}

@@

We examine the relationship between $p_\mathrm{in/out}$ and $c_\mathrm{in}$ by constructing the two-dimensional kernel density estimation (KDE) plots seen in Figure \ref{fig:kde}.
The KDE plots allow us to view the measured distributions of $p_\mathrm{in/out}$ and $c_\mathrm{in}$, and develop a visual impression of how well these distributions correlate with one another.
For this analysis we considered two VI sites, NAS North Island and the ASU House.
The ASU House dataset was divided into two periods, one before and the other after the land drain (called the preferential pathway (PP) from  here on) had been closed.
By comparing these two periods on a single plot, the impact of the preferential pathway becomes clearer.\par

@@

The above was proven by a second simulation, where the model was rerun with the preferential pathway present, but with the permeable (gravel) layer in the subslab removed and replaced by the surrounding soil (sandy loam).
This gave a ”uniform soil” scenario the results of which are shown as an orange line in the top panel of Figure \ref{fig:land_drain_scenarios}.
This simulation demonstrates that without a permeable subslab to effectively allow the ”advective potential” to be realized, existence of preferential pathway will actually not impact a VI site very much.
In order for a preferential pathway to significantly contribute to VI, this requires a scenario involving good advective communication between it the indoor environment.
These requirements were met at the ASU House.\par

@@

The model is also able to capture the weak trend in $\alpha_\mathrm{gw}$ with $p_\mathrm{in/out}$ when a preferential pathway is absent, but when there still exists a permeable subslab region.
These results are shown in the bottom panel of Figure \ref{fig:land_drain_scenarios}.  
These results are again in agreement with what was observed at the ASU House when the preferential pathway was closed, i.e. that there was a much more modest variation in indoor air concentration, irrespective of pressure, when the preferential pathway was cut off.\par

@@

@@

\begin{figure}
 \caption{Transient simulation of a "typical" VI day, using diurnal indoor/outdoor pressure difference and air exchange rate as inputs. Effect of preferential pathway considered.}\label{fig:diurnal}
 \includegraphics[width=\textwidth]{diurnal.pdf}
\end{figure}

The result of these simulations are shown in the bottom two panels of Figure \ref{fig:diurnal}, where the left and right panels show the results of open and closed preferential pathways, respectively.
The "max change" value in the legends is the quotient of the lowest and highest predicted concentrations, i.e. a value of two indicate that the maximum daily concentration is twice as high as the lowest.
This quantity may be compared with the value that is plotted for "one day" in Figure \ref{fig:resampling}.
When the preferential pathway is open, there is a maximum daily variation of roughly a factor of 5, irrespective of whether $A_e$ fluctuates or not, which is somewhat more than the maximum daily variation shown in Figure \ref{fig:resampling}.
The relatively small difference between the variable and constant $A_e$ cases indicates that most of the variability during a "typical" day is here attributable to fluctuations in $p_\mathrm{in/out}$, i.e. the contaminant transport into the modeled structure is advection dominated.
Even for the small fluctuations in $p_\mathrm{in/out}$ the contaminant entry rate fluctuation drives the observed indoor concentration.
When the preferential pathway is closed the story is quite different.
When air exchange rate is held constant, there is essentially no variation in $c_\mathrm{in}$.
This is again not surprising, as Figure \ref{fig:land_drain_scenarios} demonstrated that when the preferential pathway is closed, the influence of $p_\mathrm{in/out}$ on contaminant entry rate (and subsequently $c_\mathrm{in}$) is small.
Combined with the small $p_\mathrm{in/out}$ this indicates that the contaminant transport into the modeled structure in this scenario is dominated by diffusion.
When the air exchange rate is allowed to fluctuate, the maximum daily variation in $c_\mathrm{in}$ is 1.68, which is in line with what is shown in Figure \ref{fig:resampling}.
This shows that for a "typical" day, when the preferential pathway is closed off, much of the daily variation in $c_\mathrm{in}$ is due to daily fluctuations in air exchange rate.\par

These results demonstrate the complicated nature of temporal variability in $c_\mathrm{in}$.
It is important to recall that only the effects of indoor/outdoor pressure difference and air exchange rate have been considered here, but slower processes, e.g. changes in groundwater contaminant concentration or various seasonal effects can also have a significant impact on VI over time.
For the shorter time periods of concern in recent studies of temporal variability in indoor contaminant concentrations we believe that these are dominated by combinations of indoor/outdoor pressure differentials and air exchange rate.
For a site where advective communication between the subsurface and the indoor is good, $p_\mathrm{in/out}$ is likely a significant determinant of $c_\mathrm{in}$ and its temporal variability.
We have shown that that such a scenario may arise due to a preferential pathway entering a permeable sub-base, but may also exist even in the absence of a preferential pathway just as the results from NAS North Island demonstrate.
At sites where advective transport into the structure is limited, much of the temporal variability in $c_\mathrm{in}$ may be attributed to natural fluctuations in air exchange rate.\par

### 2019-04-15 tex/supporting-information/esthag_supporting.tex
Removed Guo (for now)
Implemented Yuanmings suggestions (among other)
Wrote abstract and references new figure directory
Last changes before submission

% additional packages
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{subcaption}
\usepackage{array,multirow,graphicx}
\graphicspath{{../../figures/2d_kde/}}
% macros

% authors
\author{Jonathan G. V. Ström}
\affiliation[Brown University]{Brown University, School of Engineering, Providence, RI, USA}
\author{Yijun Yao}
\affiliation[Brown University]{Brown University, School of Engineering, Providence, RI, USA}
\author{Yuanming Guo}
\affiliation[Arizona State University]{Arizona State University}
\author{Eric M. Suuberg}
\email{eric_suuberg@brown.edu}
\affiliation[Brown University]{Brown University, School of Engineering, Providence, RI, USA}

@@

% keywords
\abbreviations{VI}
\keywords{Vapor intrusion, Preferential pathways}

@@

\newpage
\subsection{Pressure and Air Exchange Rate}
\begin{figure}
  \caption{2D KDE figure showing distributions and relationship between indoor/outdoor pressure difference and air exchange rate. The seasonal median $p_\mathrm{in/out}$ and $A_e$ shown. Only non-CPM period considered.}
  \includegraphics[width=\textwidth]{pressure_air_exchange_rate.pdf}
\end{figure}

>>>>

% additional packages
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{subcaption}
\usepackage{array,multirow,graphicx}
\graphicspath{{./../figures/}}
% macros

% authors
\author{Jonathan G. V. Ström}
\affiliation[Brown University]{Brown University, School of Engineering, Providence, RI, USA}
\author{Yuanming Guo}
\affiliation[Arizona State University]{Arizona State University, School of Sustainable Engineering and the Building Environment, Tempe, AZ, USA}
\author{Yijun Yao}
\affiliation[Brown University]{Brown University, School of Engineering, Providence, RI, USA}
\author{Eric M. Suuberg}
\email{eric_suuberg@brown.edu}
\affiliation[Brown University]{Brown University, School of Engineering, Providence, RI, USA}

@@

% keywords
\abbreviations{VI}
\keywords{Vapor intrusion, Preferential pathways, Temporal variability, Finite element modeling, Air exchange rate, Indoor/outdoor pressure difference}

@@

\newpage
\subsection{Pressure and Air Exchange Rate}
\begin{figure}
  \caption{2D KDE figure showing distributions and relationship between indoor/outdoor pressure difference and air exchange rate. The seasonal median $p_\mathrm{in/out}$ and $A_e$ are indicated by the location of the respective arrow tips. Only non-CPM period considered.}
  \includegraphics[width=\textwidth]{pressure_air_exchange_rate.pdf}
\end{figure}
